---
title: "Make spillover matrix for compensation in IMC image"
output: html_notebook
---

```{r Libraries}
library(CATALYST)
library("EBImage")
library(data.table)
library(fields)
library(ggplot2)
library(flowCore)
library(plyr)
library(dplyr)
library(dtplyr)
library(MASS)
library(viridis)
library(RColorBrewer)
library(raster)
library(stringr)

```


Point to the singlestain folders
```{r Setup}
fol_ss1 = c('/home/jana/Desktop/compensation_matrix/2017-09-22/')
fol_ss2 = c('/home/jana/Desktop/compensation_matrix/2017-09-12/')
fol_ss3 = c('/home/jana/Desktop/compensation_matrix/2017-11-15/')
fol_out = '/home/jana/'
prefix = '2018_05_03'

```


Load single stains
```{r Load from Folders or Zips}

load_ss_zip <- function(fol_ss3){
  fns_inzip <- unzip(fol_ss3, list=T) %>%
            do(as.data.frame(.$Name[ endsWith(.$Name, '.txt')]))

  fns_inzip = fns_inzip[,1]
  imgs.ss = lapply(fns_inzip, function(x){
              fread(paste0('unzip -qp ', fol_ss3, ' ', gsub(" ", '\\\\ ', x)))
                })
  
  names(imgs.ss) <- fns_inzip
return(imgs.ss)
}

load_ss_fol <- function(fol_ss3){
  fns_txt <- list.files(fol_ss3,pattern = '*.[0-9]+.txt$') 
  imgs.ss <- lapply(fns_txt, function(x){
              fread(file.path(fol_ss3, x))})
  names(imgs.ss) <- fns_txt
  return(imgs.ss)
}
list_img_ss <-lapply(fol_ss3, load_ss_fol)
names(list_img_ss) <- fol_ss3
#imgs.ss = do.call(list, unlist(imgs.ss, recursive=FALSE))
```

Fix the names within the text files to by CATALYST compatible
```{r}
fixnames <- function(imgdat){
    imgdat = copy(imgdat)
    dat =imgdat
    
    colnames(dat) = sapply(colnames(dat), function(x) gsub('.*\\(', '',x))
    colnames(dat) = sapply(colnames(dat), function(x) gsub('\\)', '',x))  
    return(dat)
}

imglist2dat <- function(datlist){
  imgdat <- rbindlist(lapply(datlist, fixnames), fill=T, idcol = 'file')
  #imgdat <- imgdat
  return(imgdat)
}

dats_raw = lapply(list_img_ss, imglist2dat)
```


Extract BC masses from the filenames
```{r Get bc masses}

for (dat in dats_raw){
  dat[, metal:= strsplit(.BY[[1]], '_')[[1]][3],by=file]
  dat[, mass:= as.numeric(str_extract_all(.BY[[1]], "[0-9]+")[[1]]),by=metal]
}
```

Look at the data
```{r}
summarize_dat <- function(dat){
  tdat = dat %>%  
    dplyr::select(-c(Start_push, End_push, Pushes_duration,   X , Y  ,  Z)) %>%
  melt.data.table(id.vars = c('metal', 'mass','file')) %>%
  do(data.table(.)[, list(med=median(value)), by=.(variable, metal, mass, file)]) 
  return(tdat)
  
}

dats_raw_sum = rbindlist(lapply(dats_raw, summarize_dat),idcol = T)
```

Look at the data
```{r fig.width=17, fig.height=10}
dats_raw_sum %>%
  ggplot(aes(x=mass, y=med, color=.id))+
  facet_wrap(~file+metal, scales = 'free_y')+
  geom_label(aes(label=variable))
  #geom_jitter(height = 0)
  
```

Depending on the intensity it could be worth to bin some (consecuteive) pixels to get a better accuracy for the estimation
```{r}

npixelbin = 3

get_consecutive_bin <- function(nel, nbin){
  idx = rep(1:ceiling(nel/nbin), each=nbin)
  return(idx[1:nel])
}

aggregate_pixels <- function(dat, n){
  tdat = dat[, rowsum(.SD, get_consecutive_bin(.N, n)) ,by=.(file, mass, metal)]
  return(tdat)
}

dats_agg <- lapply(dats_raw, function(x) aggregate_pixels(x, n=npixelbin))
dats_agg_sum = rbindlist(lapply(dats_agg, summarize_dat), idcol = T)
```

After aggregation
```{r fig.width=17, fig.height=10}
dats_agg_sum %>%
  ggplot(aes(x=mass, y=med, color=.id))+
  facet_wrap(~file+metal, scales = 'free_y')+
  geom_label(aes(label=variable))
  #geom_jitter(height = 0)
  
```

Define helper funcitons
```{r Define helper functions}
filter_rare_bc <- function(re, minevents){
  stats = table(re@bc_ids)
  nonfreq = names(stats)[stats <minevents]
  re@bc_ids[re@bc_ids %in% nonfreq] = '0'
  return(re)
  
}

ensure_correct_bc <- function(re, mass){
  re@bc_ids[re@bc_ids != as.character(mass)] = '0'
  return(re)
}


re_from_dat <- function(dat, bc_ms, minevents=10){
  ff = dat %>%
      dplyr::select(-c(file, mass, metal)) %>%
      as.matrix.data.frame() %>%
      flowFrame()
  
  
  re <- CATALYST::assignPrelim(x=ff,y= bc_ms)
  re <- estCutoffs(re)
  re <- applyCutoffs(re)

  # filter for conditions with less then minevents
  #re = ensure_correct_bc(re, dat[, mass])
  re = filter_rare_bc(re, minevents)
  return(re)
}



sm_from_re <- function(re, ...){
  sm = computeSpillmat(re, ...)
  return(sm)
}

sm_from_dat <- function(dat, bc_ms, minevents=10, ...){
  re <- re_from_dat(dat, bc_ms, minevents)
  sm <- sm_from_re(re, ...)
  return(sm)
}
```

Write out spillover matrix
```{r Binned}

res = lapply(dats_agg, function(x) re_from_dat(x, x[!is.na(mass), unique(mass)]))
sms = lapply(res, function(x) sm_from_re(x))
sm = data.frame(sms[[1]])
fwrite(sm, file='/home/jana/sm_test.csv', col.names = TRUE, row.names = TRUE)
```

Plot spillover matrix
```{r}
i = 1
bc_ms =  dats_agg[[i]][!is.na(mass), unique(mass)]
custom_isotope_list <- c(CATALYST::isotope_list, list(BCKG=190, ArAr=80))
CATALYST::plotSpillmat(bc_ms,sms[[1]], isotope_list=custom_isotope_list)
```

```{r quick check if the debarcoding went well, so all the events were correclty classified}
i =1
re = res[[i]]
dat = copy(dats_agg[[i]])
dat[ , bc := bc_ids(re)]

dat[bc!= '0', sum(as.character(mass) != bc), by=metal]
```